---
title       : Inter-Rater Reliability
subtitle    : 
date        : "`r Sys.time()`"
author      : Mark White
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]     # {quiz, bootstrap}
mode        : standalone # {selfcontained, draft}
knit        : slidify::knit2slides
---

```{r setup, include=F}
require(slidify)
# setwd("r/IRR")
# rm(list=ls())
source("../../../../Research/R libraries/DefaultScriptMark.R")
require(ggplot2)
require(data.table)
require(lme4)
require(gdata)
require(ordinal)
require(plyr)
require(xtable)
options(digits=2)
options(xtable.type="html",xtable.include.rownames=F)
knitr::opts_chunk$set(echo = F)


raters<-c("303","315","318","319","322","342","343","344","366","379","383","398")
items  <-c("FFT_RR_1","FFT_CL_1","FFT_MCP_1","FFT_MSB_1","FFT_OPS_1","FFT_CS_1" ,"FFT_KC_1" ,"FFT_QDT_1","FFT_ESL_1","FFT_UAI_1","FFT_FR_1" )
```
```{r restructure, include=F,cache=T}
load(file=file.path("C:/Users/mrkwht/Documents/UTQ/UTQProject/data/Compiled_ObsData.rData"))

Video[,matchcols(Video,c("MQI|CLASS|IDAP|Sec|PreObsQSummary|PLATO"),c("DOMAIN","STRUCT","READ","SectionID")):=NULL]
Video[,c(matchcols(Video,"TKAS|CameraFile|Just"),"ObsPatt","WorkingNotes"):=NULL]
dim(FFT <- Video[,.SD,.SDcols=c("SectionID","TeacherID","SchoolID","LessonID","DistrictID","ScoreType","ScoreSessionID","Subject","Grade","FFT_ObserverID","FFT_StartTime","FFT_NoteEndTime","FFT_EndTime","FFT_TimestampRevised","FFT_RR_1","FFT_CL_1","FFT_MCP_1","FFT_MSB_1","FFT_OPS_1","FFT_CS_1","FFT_KC_1","FFT_QDT_1","FFT_ESL_1","FFT_UAI_1","FFT_FR_1","PLATO_DOMAIN_Read_1","PLATO_DOMAIN_Read_2","PLATO_DOMAIN_Writ_1","PLATO_DOMAIN_Writ_2","PLATO_DOMAIN_Lit_1","PLATO_DOMAIN_Lit_2","PLATO_DOMAIN_Oral_1","PLATO_DOMAIN_Oral_2","PLATO_DOMAIN_Vocab_1","PLATO_DOMAIN_Vocab_2","PLATO_DOMAIN_Grammar_1","PLATO_DOMAIN_Grammar_2","PLATO_DOMAIN_Research_1","PLATO_DOMAIN_Research_2","PLATO_STRUCT_Tlecture_1","PLATO_STRUCT_Tlecture_2","PLATO_STRUCT_ShortSResp_1","PLATO_STRUCT_ShortSResp_2","PLATO_STRUCT_SmGrpUnstr_1","PLATO_STRUCT_SmGrpUnstr_2","PLATO_STRUCT_SmGrpStr_1","PLATO_STRUCT_SmGrpStr_2","PLATO_STRUCT_WholeGrp_1","PLATO_STRUCT_WholeGrp_2","PLATO_STRUCT_StdPres_1","PLATO_STRUCT_StdPres_2","PLATO_STRUCT_IndpWork_1","PLATO_STRUCT_IndpWork_2","PLATO_STRUCT_UseSPrLang_1","PLATO_STRUCT_UseSPrLang_2","PLATO_STRUCT_Differentiated_1","PLATO_STRUCT_Differentiated_2","PLATO_READ_Independent_1","PLATO_READ_Independent_2","PLATO_READ_Timed_1","PLATO_READ_Timed_2","PLATO_READ_Whole_1","PLATO_READ_Whole_2","PLATO_READ_Small_1","PLATO_READ_Small_2","Semester","ObsDate","ObsTime","CAM_TViewNoAudio","CAM_SViewNoAudio","CAM_VideoQuality","CAM_Adults","QC_TVIEW_VID_CLARITY","QC_TVIEW_VID_OPERATION","QC_TVIEW_TAUDIO","QC_TVIEW_SAUDIO","QC_SVIEW_VID_CLARITY","QC_SVIEW_VID_OPERATION","QC_SVIEW_TAUDIO","QC_SVIEW_SAUDIO","QC_SYNC","QC_SUMMARY_SCORE","CameraOperator","ClassLength","TCH_GENDER","TCH_RACE","TCH_YRS_MID","TCH_SUB_K12","TCH_ALT_CERT","TCH_TEAM_TCH","TCH_INTELL_CNTCHG","TCH_INTELL_AMT","TCH_INTELL_LRN_NEW","TCH_INTELL_CHG_LOT","TCH_INTELL_CNCHG","TCH_INTELL_CHG_AMT","VAM_T_Reg6","VAM_T_Reg6_AltYr","NumSt" )])
dim(FFT<-FFT[ScoreType!="Live",])

FFT[,VidQuality:=apply(data.frame(QC_TVIEW_VID_CLARITY,QC_TVIEW_VID_OPERATION,QC_SVIEW_VID_CLARITY,QC_SVIEW_VID_OPERATION),1,mean,na.rm=T)]
FFT[,c("QC_TVIEW_VID_CLARITY","QC_TVIEW_VID_OPERATION","QC_SVIEW_VID_CLARITY","QC_SVIEW_VID_OPERATION"):=NULL]

FFT[,PLATO_DOMAIN_Read          :=(PLATO_DOMAIN_Read_1          +PLATO_DOMAIN_Read_2          )/2]
FFT[,PLATO_DOMAIN_Writ          :=(PLATO_DOMAIN_Writ_1          +PLATO_DOMAIN_Writ_2          )/2]
FFT[,PLATO_DOMAIN_Lit           :=(PLATO_DOMAIN_Lit_1           +PLATO_DOMAIN_Lit_2           )/2]
FFT[,PLATO_DOMAIN_Oral          :=(PLATO_DOMAIN_Oral_1          +PLATO_DOMAIN_Oral_2          )/2]
FFT[,PLATO_DOMAIN_Vocab         :=(PLATO_DOMAIN_Vocab_1         +PLATO_DOMAIN_Vocab_2         )/2]
FFT[,PLATO_DOMAIN_Grammar       :=(PLATO_DOMAIN_Grammar_1       +PLATO_DOMAIN_Grammar_2       )/2]
FFT[,PLATO_DOMAIN_Research      :=(PLATO_DOMAIN_Research_1      +PLATO_DOMAIN_Research_2      )/2]
FFT[,PLATO_STRUCT_Tlecture      :=(PLATO_STRUCT_Tlecture_1      +PLATO_STRUCT_Tlecture_2      )/2]
FFT[,PLATO_STRUCT_ShortSResp    :=(PLATO_STRUCT_ShortSResp_1    +PLATO_STRUCT_ShortSResp_2    )/2]
FFT[,PLATO_STRUCT_SmGrpUnstr    :=(PLATO_STRUCT_SmGrpUnstr_1    +PLATO_STRUCT_SmGrpUnstr_2    )/2]
FFT[,PLATO_STRUCT_SmGrpStr      :=(PLATO_STRUCT_SmGrpStr_1      +PLATO_STRUCT_SmGrpStr_2      )/2]
FFT[,PLATO_STRUCT_WholeGrp      :=(PLATO_STRUCT_WholeGrp_1      +PLATO_STRUCT_WholeGrp_2      )/2]
FFT[,PLATO_STRUCT_StdPres       :=(PLATO_STRUCT_StdPres_1       +PLATO_STRUCT_StdPres_2       )/2]
FFT[,PLATO_STRUCT_IndpWork      :=(PLATO_STRUCT_IndpWork_1      +PLATO_STRUCT_IndpWork_2      )/2]
FFT[,PLATO_STRUCT_UseSPrLang    :=(PLATO_STRUCT_UseSPrLang_1    +PLATO_STRUCT_UseSPrLang_2    )/2]
FFT[,PLATO_STRUCT_Differentiated:=(PLATO_STRUCT_Differentiated_1+PLATO_STRUCT_Differentiated_2)/2]
FFT[,PLATO_READ_Independent     :=(PLATO_READ_Independent_1     +PLATO_READ_Independent_2     )/2]
FFT[,PLATO_READ_Timed           :=(PLATO_READ_Timed_1           +PLATO_READ_Timed_2           )/2]
FFT[,PLATO_READ_Whole           :=(PLATO_READ_Whole_1           +PLATO_READ_Whole_2           )/2]
FFT[,PLATO_READ_Small           :=(PLATO_READ_Small_1           +PLATO_READ_Small_2           )/2]
FFT[,c("PLATO_DOMAIN_Read_1","PLATO_DOMAIN_Read_2","PLATO_DOMAIN_Writ_1","PLATO_DOMAIN_Writ_2","PLATO_DOMAIN_Lit_1","PLATO_DOMAIN_Lit_2","PLATO_DOMAIN_Oral_1","PLATO_DOMAIN_Oral_2","PLATO_DOMAIN_Vocab_1","PLATO_DOMAIN_Vocab_2","PLATO_DOMAIN_Grammar_1","PLATO_DOMAIN_Grammar_2","PLATO_DOMAIN_Research_1","PLATO_DOMAIN_Research_2","PLATO_STRUCT_Tlecture_1","PLATO_STRUCT_Tlecture_2","PLATO_STRUCT_ShortSResp_1","PLATO_STRUCT_ShortSResp_2","PLATO_STRUCT_SmGrpUnstr_1","PLATO_STRUCT_SmGrpUnstr_2","PLATO_STRUCT_SmGrpStr_1","PLATO_STRUCT_SmGrpStr_2","PLATO_STRUCT_WholeGrp_1","PLATO_STRUCT_WholeGrp_2","PLATO_STRUCT_StdPres_1","PLATO_STRUCT_StdPres_2","PLATO_STRUCT_IndpWork_1","PLATO_STRUCT_IndpWork_2","PLATO_STRUCT_UseSPrLang_1","PLATO_STRUCT_UseSPrLang_2","PLATO_STRUCT_Differentiated_1","PLATO_STRUCT_Differentiated_2","PLATO_READ_Independent_1","PLATO_READ_Independent_2","PLATO_READ_Timed_1","PLATO_READ_Timed_2","PLATO_READ_Whole_1","PLATO_READ_Whole_2","PLATO_READ_Small_1","PLATO_READ_Small_2"):=NULL]
FFT[,TCH_INTELL:=apply(.SD,1,mean,na.rm=T),.SDcols=matchcols(FFT,"TCH_INTELL")]
FFT[,matchcols(FFT,"TCH_INTELL_"):=NULL]

FFT[,matchcols(FFT,"PLATO"):=lapply(.SD,function(x) mean(unique(na.omit(x)))),by=LessonID,.SDcols=matchcols(FFT,"PLATO")]

dim(FFT<-FFT[!is.na(FFT_ObserverID),])
DoubleLesson <-FFT[ScoreType=="Double",unique(LessonID)]
dim(FFT<-FFT[LessonID %in% DoubleLesson,])

FFTtall=melt.data.table(FFT,measure.vars=matchcols(FFT,"1$",c("NumNotes","Just")))
FFTtall[,variable:=factor(sub("FFT_","",sub("_1","",char(variable))))]
FFTwide<-dcast.data.table(FFTtall,LessonID+SectionID+TeacherID+SchoolID+DistrictID+Subject+Grade~variable+FFT_ObserverID)
FFTtall[,valueM:=mean(value),by=c("LessonID","variable")]
```

## Inter-Rater Reliability/Agreement
<br> <br>
- Goal of measurement is to capture construct
- Specific rater is rarely part of context
- Two questions arise
  - Are measurements accurately capturing construct?
  - How much error is being introduced by raters?
- <http://www.john-uebersax.com/stat/agree.htm>

---
## Different Types of Inter-Rater Reliability
<br> <br>
- Agreement versus Reliability
- True score available or not

--- &twocol
## Which Scenario has Better Raters?

***=left
<br> <br>
<div style="text-align:center;width:100%">
<table style="width:50%;margin-left:auto;margin-right:auto">
  <tr>
    <th></th> <th></th> <th colspan="2">Rater 1</th>
  </tr>
  <tr> <td rowspan="3">Rater 2</td> 
       <td> </td> <td><b>0</b></td> <td><b>1</b></td> </tr>
  <tr> <td>0</td> <td>40</td> <td>10</td> </tr>
  <tr> <td>1</td> <td>10</td> <td>40</td> </tr>
</table>
</div>

<br> <br>
<div style="text-align:center;width:100%">
<table style="width:50%;margin-left:auto;margin-right:auto">
  <tr>
    <th></th> <th></th> <th colspan="2">Rater 1</th>
  </tr>
  <tr> <td rowspan="3">Rater 2</td> 
       <td> </td> <td><b>0</b></td> <td><b>1</b></td> </tr>
  <tr> <td>0</td> <td>40</td> <td>20</td> </tr>
  <tr> <td>1</td> <td>0</td> <td>40</td> </tr>
</table>
</div>

***=right
<br> <br>
<div style="text-align:center;width:100%">
<table style="width:50%;margin-left:auto;margin-right:auto">
  <tr>
    <th></th> <th></th> <th colspan="2">Rater 1</th>
  </tr>
  <tr> <td rowspan="3">Rater 2</td> 
       <td> </td> <td><b>0</b></td> <td><b>1</b></td> </tr>
  <tr> <td>0</td> <td>5</td> <td>10</td> </tr>
  <tr> <td>1</td> <td>10</td> <td>75</td> </tr>
</table>
</div>

<br> <br>
<div style="text-align:center;width:100%">
<table style="width:50%;margin-left:auto;margin-right:auto">
  <tr>
    <th></th> <th></th> <th colspan="2">Rater 1</th>
  </tr>
  <tr> <td rowspan="3">Rater 2</td> 
       <td> </td> <td><b>0</b></td> <td><b>1</b></td> </tr>
  <tr> <td>0</td> <td>5</td> <td>18</td> </tr>
  <tr> <td>1</td> <td>2</td> <td>75</td> </tr>
</table>
</div>

---
## Chance Corrected Agreement
<br>
- Raters may be correct simply due to chance
- The distribution of scores will affect chance agreement
- How to adjust percent agreement for guessing
  - Kappa
  - Ordinal Kappa
  - PABAK

--- &twocol
##  Kappa 

***=left
- Assume Guessing Distribution Matches Observed Scores
- Assume Scoring Independent
- $P(R_1\bigcap R_2=1)= (b+d) * (c+d)$
- $P(R_1\bigcap R_2=0)= (a+c) * (a+b)$
- $P_C(R_1=R_2 by Chance)=$   
      $\quad P(R_1\bigcap R_2=1)+P(R_1\bigcap R_2=0)$
- $P_O=a+d$
- $Kappa= \frac{P_O-P_C}{1-P_C}$

***=right
<br> <br>
<div style="text-align:center;width:100%">
<table style="width:50%;margin-left:auto;margin-right:auto">
  <tr>
    <th></th> <th></th> <th colspan="2">Rater 1</th>
  </tr>
  <tr> <td rowspan="3">Rater 2</td> 
       <td> </td> <td><b>0</b></td> <td><b>1</b></td> </tr>
  <tr> <td>0</td> <td>a</td> <td>b</td> </tr>
  <tr> <td>1</td> <td>c</td> <td>d</td> </tr>
</table>
</div>

---
##  Kappa 
<br>
- Prevalence - distribution of event in sample
- Bias - Rater disagreement over prevalence
- Both still impact scores 
  - Implies Kappa is sample dependent
- Interpretation is not clear
- Many suggest not using standard rules of thumb

***=footnotes
Cicchetti & Feinstein. (1990). High Agreement but Low Kappa: Resolving the Paradoxes. Journal of Clinical Epidemiology

--- .compact
##  Kappa Extensions

- Weighted Kappa - for ordinal/interval data
  - Weights disagreement based on distance between rated values
  - Choice of weights is arbitrary
- Scott's PI (Intraclass Kappa)
  - Uses overall distribution for guessing parameter
  - Can be used as multi-rater Kappa
- Prevalence and Bias Adjusted Kappa (PABAK)
  - $(k*P_O-1)/(k-1)$
  - Assumes away prevalence and bias problems
- Krippendorf Alpha
  - Alternative that generalizes to ICC
  - Multi-rater, nominal/ordinal/interval measures
  - Unbalanced designs

---
##  Theoretical Issues: Kappa
<br>
- Kappa is better than simple percent agreement, but 
  - picks arbitrary way of adjusting for chance
    - Raters likely do not always guess based on marginal frequency
  - Prevalence and Bias still raise issues and cause paradoxes

--- &twocol
## Specific Agreement

***=left
- Generalization of Specificity Sensitivity
  - http://www.john-uebersax.com/stat/agree.htm
  - Cicchetti & Feinstein (1990)
- Represents Ratings with 1 value per outcome possibility
  - $P(R_1=R_2|R_1 or R_2=1)=\frac{d+d}{(b+d)+(c+d)}$
  - $P(R_1=R_2|R_1 or R_2=0)=\frac{a+a}{(b+a)+(c+a)}$

***=right
<br> <br>
<div style="text-align:center;width:100%">
<table style="width:50%;margin-left:auto;margin-right:auto">
  <tr>
    <th></th> <th></th> <th colspan="2">Rater 1</th>
  </tr>
  <tr> <td rowspan="3">Rater 2</td> 
       <td> </td> <td><b>0</b></td> <td><b>1</b></td> </tr>
  <tr> <td>0</td> <td>a</td> <td>b</td> </tr>
  <tr> <td>1</td> <td>c</td> <td>d</td> </tr>
</table>
</div>

---
## Inter-Rater Reliability
<br>
- Correlation
- Polychoric Correlation
- ICC
- G Theory

---
## Intra-Class Correlation 
- $\sigma_i^2$ item RE, $\sigma_r^2$ rater RE, $\sigma_e^2$ residual error
- ICC(1) - $\sigma_i^2/(\sigma_i^2+\sigma_e^2)$
  - Raters nested in Items
- Random Rater:
  - ICC(C,1) - $\sigma_i^2/(\sigma_i^2+\sigma_{ir}^2+\sigma_e^2)$
  - ICC(A,1) - $\sigma_i^2/(\sigma_i^2+\sigma_r^2+\sigma_{ir}^2+\sigma_e^2)$
  - ICC(C,k) - $\sigma_i^2/(\sigma_i^2+\sigma_{ir}^2/k+\sigma_e^2/k)$
- Fixed Rater: 
  - ICC(C,1) - $(\sigma_i^2+\sigma_{ir}^2/(k-1))/(\sigma_i^2+\sigma_{ir}^2+\sigma_e^2)$
  - ICC(A,1) - $(\sigma_i^2+\sigma_{ir}^2/(k-1))/(\sigma_i^2+\sigma_r^2+\sigma_{ir}^2+\sigma_e^2)$
  - ICC(C,k) - $(\sigma_i^2+\sigma_{ir}^2/(k-1))/(\sigma_i^2+\sigma_{ir}^2/k+\sigma_e^2/k)$

***=footnotes
McGraw & Wong. (1996). Forming inferences about some intraclass correlation coefficients. Psychological methods.

---
## G Theory
<br>
- Estimates all variance components 
- Allows combination of variance components into various ICC measures

---
## Examples from FFT in UTQ
```{r echo=F,results="asis"}
print(xtable(dcast.data.table(FFTtall,LessonID+variable~ScoreType,value.var="FFT_ObserverID")[variable=="RR",table(Regular,Double)]),include.rownames=T)
```

---
## Examples from FFT in UTQ

```{r include=F, cache=T}
gt<-list()
for (i in unique(FFTtall$variable))
  gt[[i]] <- lmer(value~1+(1|LessonID)+(1|FFT_ObserverID),FFTtall,subset=variable==i)
VarPer <- function(x) { 
  x <- summary(x)$varcor
  out <- c(unlist(x),Res=attr(x,"sc")^2)
  round(out / sum(out)*100,1)
}
gtpct<-t(sapply(gt,VarPer))
gtpct<-data.frame(variable=rownames(gtpct),gtpct)


MabsD<-FFTtall[,mean(abs(value-valueM)),by=c("FFT_ObserverID","variable")][,xtabs(V1~FFT_ObserverID+ variable)]
RMSE <-FFTtall[,sd(value-valueM),by=c("FFT_ObserverID","variable")][,xtabs(V1~FFT_ObserverID+ variable)]

kappa.pair   <- data.frame(variable=unique(FFTtall$variable),kappa.pair=sapply(unique(FFTtall$variable), function(x) mean(apply(combinat::combn(matchcols(FFTwide,x),2),2, function(y) { irr::kappa2(as.data.frame(FFTwide)[,y],weight = "squared")$value }),na.rm=T)))
kappa.ignore <- dcast(FFTtall,LessonID+variable~ScoreType)[,data.frame(Kappa=irr::kappa2(data.frame(Regular,Double),weight="squared")$value),by=variable]

ExpProb <- FFTtall[,sum(prop.table(base::table(value))^2),by=variable]
ScottPi <- function(x,i) {
  tab <- base::table(x)
  n<-sum(tab)
  agree<-sum(diag(tab))
  Po <- agree/n
  Pe <-ExpProb[variable==i,V1]
  return( (Po-Pe)/(1-Pe) )
}
PABAK <- function(x) (4*sum(diag(prop.table(base::table(x))))-1)/3

pabak<-dcast(FFTtall,LessonID+variable~ScoreType)[,PABAK(data.frame(Regular,Double)),by=variable]
pi.pair   <- data.frame(variable=unique(FFTtall$variable),pi.pair=sapply(unique(FFTtall$variable), function(x) mean(apply(combinat::combn(matchcols(FFTwide,x),2),2, function(y) { ScottPi(as.data.frame(FFTwide)[,y],x) }),na.rm=T)))
pi.ignore <- dcast(FFTtall,LessonID+variable~ScoreType)[,data.frame(pi=ScottPi(data.frame(Regular,Double),variable)),by=variable]

ka <- data.frame(variable=unique(FFTtall$variable),alpha=sapply(unique(FFTtall$variable), function(x) irr::kripp.alpha(data.matrix(data.frame(FFTwide)[,matchcols(FFTwide,x)]),method="ordinal")$value))

tab <- data.table(variable=names(apply(MabsD,2,mean)),
             AbsD=apply(MabsD,2,mean),
             RMSE=apply(RMSE,2,mean))
names(pabak)<-c("variable","PABAK")
tab<-merge(tab,merge(kappa.pair,kappa.ignore,by="variable",all=T),by="variable",all=T)
tab<-merge(kappa.pair,kappa.ignore,by="variable",all=T)
tab<-merge(tab,pabak,by="variable",all=T)
tab<-merge(tab,pi.pair,by="variable",all=T)
tab<-merge(tab,pi.ignore,by="variable",all=T)
tab<-merge(tab,ka,by="variable",all=T)
tab<-merge(tab,gtpct,by="variable",all=T)
```
```{r echo=F,results="asis"}
addtorow <-list()
addtorow$pos <- list(0)
addtorow$command <- c(paste0("<col>",
                             "<colgroup span='2'></colgroup>","<col>",
                             "<colgroup span='2'></colgroup>","<col>",
                             "<colgroup span='3'></colgroup>",
                             "<tr>","<th rowspan='2' style='border-right:1pt solid black;vertical-align:middle'>Dimension</th>",
                             "<th colspan='2' scope='colgroup' style='border-right:1pt solid black; text-align:center'>Kappa</th>",
                             "<th rowspan='2' scope='col' style='border-right:1pt solid black;vertical-align:middle'>PABAK</th>",
                             "<th colspan='2' scope='colgroup' style='border-right:1pt solid black;text-align:center'>Scott's Pi</th>",
                             "<th  rowspan='2' scope='col' style='border-right:1pt solid black;text-align:center'>Kripp<br>Alpha</th>",
                             "<th colspan='3' scope='colgroup' style='border-right:0;text-align:center'>Var. Decomposition</th>",
                             "</tr>","<tr>",
                             "<th scope='col' style='border-right:0'>Pairwise</th><th scope='col' style='border-right:1pt solid black'>Incorrect</th>",
                             "<th scope='col' style='border-right:0'>Pairwise</th><th scope='col' style='border-right:1pt solid black'>Incorrect</th>",
                             "<th scope='col' style='border-right:0'>Day</th><th scope='col' style='border-right:0'>Rater</th><th scope='col' style='border-right:0'>Day*Rater</th>",
                             "</tr>",collapse=" "))

print(xtable(tab[c(5,10,1,6,7,8,2,9,3,11,4),]),html.table.attributes="class='Example'",add.to.row=addtorow,include.colnames=F)

```

---
## Implicit Assumptions Never Discussed
<br>
- Agreement is constant over full scale
<br>
- Agreement is constant over days

---
## Specific Agreement

```{r echo=F,results="asis"}
SpecAgree <- function(x,y) { 
  tab<-base::table(factor(x,levels=1:4),factor(y,levels=1:4))
  p1=2*tab[1,1]/sum(tab[1,],tab[,1])
  p2=2*tab[2,2]/sum(tab[2,],tab[,2])
  p3=2*tab[3,3]/sum(tab[3,],tab[,3])
  p4=2*tab[4,4]/sum(tab[4,],tab[,4])
  return(c(Agree=sprintf("%.0f%%", 100*sum(diag(prop.table(tab)))),
           "1"=sprintf("%.0f%% (%.1f)",100*p1,100*sqrt(p1*(1-p1)/sum(tab[1,],tab[,1]))),
           "2"=sprintf("%.0f%% (%.1f)",100*p2,100*sqrt(p2*(1-p2)/sum(tab[2,],tab[,2]))),
           "3"=sprintf("%.0f%% (%.1f)",100*p3,100*sqrt(p3*(1-p3)/sum(tab[3,],tab[,3]))),
           "4"=sprintf("%.0f%% (%.1f)",100*p4,100*sqrt(p4*(1-p4)/sum(tab[4,],tab[,4]))))) }
sagree=data.frame(variable=unique(FFTtall$variable),ldply(unique(FFTtall$variable),function(x) dcast(FFTtall,LessonID+variable~ScoreType)[variable==x,SpecAgree(Regular,Double)]))
sagree[c(1,3,4,9,10),"X4"]<-""
#
addtorow <-list()
addtorow$pos <- list(0)
addtorow$command <- c(paste0("<tr>","<th></th>","<th style='text-align:center;vertical-align:bottom'>Perc<br>Agree</th>",
                             "<th style='text-align:center;vertical-align:bottom'>1</th><th style='text-align:center;vertical-align:bottom'>2</th>","<th style='text-align:center;vertical-align:bottom'>3</th><th style='text-align:center;vertical-align:bottom'>4</th>",
                             "</tr>",collapse=" "))

print(xtable(sagree[c(7,1:6,8:11),]),html.table.attributes="class='CT'",add.to.row=addtorow,include.colnames=F)

```

---
## Are some lessons differentially hard to score?

- Could find very limited evidence of rater error systematically relating to lesson characteristics
- Strong negative relationship of disagreement and mean score on items
  - especially RR, MSB, CS
- Lessons given in May were harder to score for MSB and RR
- 

```{r diffdisagree, echo=F, include=F}

FFTtall[,scorediff:=mean(abs(value-valueM),na.rm=T),by=c("LessonID","variable")]
FFTtall[,cor(scorediff,valueM,use="pair"),by=variable]

summary(FFTlesson <- FFTtall[,unique(.SD),.SDcols=c("SectionID","TeacherID","SchoolID","LessonID","DistrictID","Subject","Grade","Semester","ObsDate","ObsTime","CAM_TViewNoAudio","CAM_SViewNoAudio","CAM_VideoQuality","CAM_Adults","QC_TVIEW_TAUDIO","QC_TVIEW_SAUDIO","QC_SVIEW_TAUDIO","QC_SVIEW_SAUDIO","QC_SYNC","QC_SUMMARY_SCORE","CameraOperator","ClassLength","TCH_GENDER","TCH_RACE","TCH_YRS_MID","TCH_SUB_K12","TCH_ALT_CERT","TCH_TEAM_TCH","VAM_T_Reg6","VAM_T_Reg6_AltYr","NumSt","VidQuality","PLATO_DOMAIN_Read","PLATO_DOMAIN_Writ","PLATO_DOMAIN_Lit","PLATO_DOMAIN_Oral","PLATO_DOMAIN_Vocab","PLATO_DOMAIN_Grammar","PLATO_DOMAIN_Research","PLATO_STRUCT_Tlecture","PLATO_STRUCT_ShortSResp","PLATO_STRUCT_SmGrpUnstr","PLATO_STRUCT_SmGrpStr","PLATO_STRUCT_WholeGrp","PLATO_STRUCT_StdPres","PLATO_STRUCT_IndpWork","PLATO_STRUCT_UseSPrLang","PLATO_STRUCT_Differentiated","PLATO_READ_Independent","PLATO_READ_Timed","PLATO_READ_Whole","PLATO_READ_Small","TCH_INTELL","variable","scorediff")])
FFTlesson<-dcast.data.table(FFTlesson,SectionID+TeacherID+SchoolID+LessonID+DistrictID+Subject+Grade+Semester+ObsDate+ObsTime+CAM_TViewNoAudio+CAM_SViewNoAudio+CAM_VideoQuality+CAM_Adults+QC_TVIEW_TAUDIO+QC_TVIEW_SAUDIO+QC_SVIEW_TAUDIO+QC_SVIEW_SAUDIO+QC_SYNC+QC_SUMMARY_SCORE+CameraOperator+ClassLength+TCH_GENDER+TCH_RACE+TCH_YRS_MID+TCH_SUB_K12+TCH_ALT_CERT+TCH_TEAM_TCH+VAM_T_Reg6+VAM_T_Reg6_AltYr+NumSt+VidQuality+PLATO_DOMAIN_Read+PLATO_DOMAIN_Writ+PLATO_DOMAIN_Lit+PLATO_DOMAIN_Oral+PLATO_DOMAIN_Vocab+PLATO_DOMAIN_Grammar+PLATO_DOMAIN_Research+PLATO_STRUCT_Tlecture+PLATO_STRUCT_ShortSResp+PLATO_STRUCT_SmGrpUnstr+PLATO_STRUCT_SmGrpStr+PLATO_STRUCT_WholeGrp+PLATO_STRUCT_StdPres+PLATO_STRUCT_IndpWork+PLATO_STRUCT_UseSPrLang+PLATO_STRUCT_Differentiated+PLATO_READ_Independent+PLATO_READ_Timed+PLATO_READ_Whole+PLATO_READ_Small+TCH_INTELL~variable,value.var="scorediff")

# hist.rug(FFTlesson$scorediff)

cor(FFTlesson[,c("PLATO_DOMAIN_Read","PLATO_DOMAIN_Writ","PLATO_DOMAIN_Lit","PLATO_DOMAIN_Oral","PLATO_DOMAIN_Vocab","PLATO_DOMAIN_Grammar","PLATO_DOMAIN_Research","PLATO_STRUCT_Tlecture","PLATO_STRUCT_ShortSResp","PLATO_STRUCT_SmGrpUnstr","PLATO_STRUCT_SmGrpStr","PLATO_STRUCT_WholeGrp","PLATO_STRUCT_StdPres","PLATO_STRUCT_IndpWork","PLATO_STRUCT_UseSPrLang","PLATO_STRUCT_Differentiated","PLATO_READ_Independent","PLATO_READ_Timed","PLATO_READ_Whole","PLATO_READ_Small"),with=F],FFTlesson[,c("CS","ESL","FR","KC","MCP","MSB","OPS","QDT","RR","CL","UAI"),with=F],use="pair")
cor(FFTlesson[,c("CAM_TViewNoAudio","CAM_SViewNoAudio","CAM_VideoQuality","CAM_Adults","QC_TVIEW_TAUDIO","QC_TVIEW_SAUDIO","QC_SVIEW_TAUDIO","QC_SVIEW_SAUDIO","QC_SYNC","QC_SUMMARY_SCORE","VidQuality","TCH_INTELL"),with=F],FFTlesson[,c("CS","ESL","FR","KC","MCP","MSB","OPS","QDT","RR","CL","UAI"),with=F],use="pair")

anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~Subject*Grade+Semester,FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~month(ObsDate)+chron::hours(ObsTime),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~factor(CameraOperator),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~cut(ClassLength,5),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~cut(TCH_YRS_MID,5),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~cut(TCH_SUB_K12,5),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~cut(NumSt,5),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~cut(VAM_T_Reg6,5),FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~TCH_GENDER+TCH_RACE+TCH_ALT_CERT,FFTlesson))
anova(lm(cbind(CS,ESL,FR,KC,MCP,MSB,OPS,QDT,RR,CL,UAI)~factor(TCH_TEAM_TCH),FFTlesson))

```
