---
title: 'Week 9: Observation Instruments'
author: "Mark White"
date: "November 12, 2018"
output: html_document
---

<style>
.forceBreak { -webkit-column-break-after: always; break-after: column; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# if (!grepl("code",getwd())) setwd("code")
require(kableExtra)
require(irr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(ltm)
require(psych)
source('../../../Research/R libraries/DefaultScriptMark.R')

filter <- dplyr::filter
rename=dplyr::rename
load("../data/Dat_Video_wHF.rData")
Videot$Observer <- relevel(Videot$Observer, "315")
Videot <- Videot %>% select(-Domain, -Interact, -ObsTime, -(CLASS_APS:PLATO_TMN), -AveSc, -DomNone, -IntNone, -PCA1, -PCA2, -Sec_PctMimp,
                            -(Sec_PctELL:Sec_PctM), -VAM_T_Reg6_AltYr_SE, -Sec_VAM_SE, -VAM_T_Reg6_SE,-Seg.Red,-Sec_PctGIFTimp,-Sec_PctAimp,
                            -Sec_PctELLimp, -SYearID, -ObsDate, -Sec_PriorAch, -ClassLength,-Sec_Cotaught,-ClassLengthc) %>%
  filter(Protocol=="FFT") %>%
  select(-CodeDate, -Protocol) 
head(Videot)
head(Videow <- Videot %>% spread(variable,value) %>% na.omit)
items <- grep("FFT_",names(Videow),value=T)
```

The code above creates two different data sets.  `Videot` stacks the variables so that each row is a response to a specific item on a specific segment.  `Videow` is wide and has each row representing a segment with scores as separate columns.

## Average

We can take the average across items to get a segment mean score and use that as the score for segment quality.

```{r}
Videow$M <- Videow %>% select(items) %>% rowMeans(na.rm=T)
hist.rug(Videow$M)
```

Testing for reliability of scores at the segment level is done with alpha

```{r}
Videow %>% select(items) %>% alpha
```

The alpha reliability, though, is only looking at the item error and other error sources remain.

## G-Theory 

For 

```{r}
# lmer(value~variable+(1|LessonID)+(1|TeacherID)+(1|SegID)+
```

## IRT



```{r}
require(mirt)

IRT1 <- mirt(select(Videow, starts_with("FFT")),itemtype="graded", model=1,
             covdata=Videow[,c("Read","Writ","Lit","Grammar","Observer")], 
             formula=~Read+Writ+Lit+Grammar+Observer)
summary(IRT1)

IRT <- mirt(select(Videow, starts_with("FFT")),itemtype="graded", model=1,
            covdata=Videow[,c("Read","Writ","Lit","Grammar","Observer","Observer")], 
            formula=~Observer)
summary(IRT)
anova(IRT,IRT1)

hist.rug(fscores(IRT1)[,1])
itemfit(IRT)
ggplot(,aes(x=round(fscores(IRT)[,1],1))) + 
  stat_smooth(aes(y=Videow$FFT_QDT,color="FFT_QDT"))+
  stat_smooth(aes(y=Videow$FFT_CL ,color="FFT_CL" ))+
  scale_color_manual("Item",
                     values=c("red","blue"))+
  labs(y="Average Value",x="Theta")

personfit(IRT) %>% .[[1]] %>% hist.rug

mlIRT <- mixedmirt(select(Videow, starts_with("FFT")),
          covdata=select(Videow, LessonID,TeacherID,Observer),
          itemtype="graded", model=1, lr.fixed=~Observer,
          lr.random= list(Teacher=~1|TeacherID)  )

coef(mlIRT) %>% .[1:11] %>% lapply(.,function(x) data.frame(x,type=row.names(x))) %>% bind_rows(.id="Item") %>% filter(type=="par")
coef(mlIRT)

str(a<-randef(mlIRT))

for (i in c("FFT_CL","FFT_CS","FFT_ESL","FFT_FR","FFT_KC","FFT_MCP","FFT_MSB","FFT_OPS","FFT_QDT","FFT_RR","FFT_UAI")) 
  print(itemplot(IRT,i, main=i))
print(itemplot(IRT,1,type="infoSE"))
itemplot(IRT,type="infotrace",item="FFT_CL")
itemplot(IRT,type="SE",item="FFT_CL")
itemplot(IRT,type="score",item="FFT_CL")

# itemplot(IRT,type="score",item="FFT_CL",shiny=T)
```


## Effects of Instructional 


```{r}
require(lme4)

summary(base<-lmer(value~variable+Observer+(1|TeacherID)+(1|LessonID),Videot))
summary(adj <-lmer(value~variable+Observer+(1|TeacherID)+(1|LessonID)+Read+Writ+Lit+Grammar+Rec+Disc+Ind,Videot))
anova(base,adj)
summary(adj2 <-lmer(value~variable+Observer+(1|TeacherID)+(1|LessonID)+DiscrIndex+Read+Writ+Lit+Grammar+Rec+Disc+Ind,Videot))

cor(data.frame(base=ranef(base)$TeacherID, adj=ranef(adj)$TeacherID, adj2=ranef(adj2)$TeacherID))


summary(adj2 <-lmer(value~variable+Observer+(1|TeacherID)+(1|LessonID)+Writ,Videot))
cor(data.frame(base=ranef(base)$TeacherID, adj=ranef(adj2)$TeacherID))
merge(ranef(base)$TeacherID %>% rownames_to_column("TID") %>% rename(Base="(Intercept)"), 
      ranef(adj2)$TeacherID %>% rownames_to_column("TID") %>% rename(adj ="(Intercept)")) %>%
  merge(Videot %>% group_by(TeacherID) %>% summarize(W=mean(Writ)) %>% ungroup %>% rename(TID="TeacherID")) %>%
  ggplot(aes(x=(Base+adj)/2,y=Base-adj,color=W)) + geom_point()
#

```



## Overall FActor Analysis

```{r, eval=F}
load("../data/Dat_Video_wHF.rData")
Vid <- Videot %>% 
  mutate(value=case_when(grepl("CLASS",variable) ~ (value-1)/2+1,
                         T ~ as.numeric(value))) %>% 
  mutate(value=resid(lm(value~variable:Observer, na.action=na.exclude))) %>%
  filter(ScoreType=="Regular") %>%
  select(SegID,variable,value) %>%
  spread(variable,value)

Vid %>% select(starts_with("PLATO"),starts_with("FFT"),starts_with("CLASS")) %>% ParallelAnalysis
Vid %>% select(starts_with("PLATO"),starts_with("FFT"),
               starts_with("CLASS")) %>% omega
Vid %>% select(starts_with("PLATO"),starts_with("FFT"),
               starts_with("CLASS")) %>% fa(6) %>% 
  print(cut=0.4)

```
