---
title: "Value Added Modeling"
author: "Mark White"
date: "December 4, 2018"
output: 
  ioslides_presentation:
     footer: "Copyright (c) 2018, Mark White"
     transition: faster
     widescreen: true
---

```{r setup, include=FALSE}
require(tidyverse)
require(knitr)
require(lme4)
require(mice)
require(kableExtra)
knitr::opts_chunk$set(echo = FALSE)
loc <- "C:/Users/mrkwht/Documents/UTQ/VAMS"

Y1 <- read_csv(file.path(loc,"StudentVAMData_Y1.csv")) %>%
  select(TID=utqid,CID=classid, dist, idc, syear, grdyr, cohort,SID=schoolid, sfrac, subject, 
         frac_t, iys_tccount, t_iyscount1, 
         eng, sped, frl, gifted, asian, black, hisp, white, male, 
         Lang=clang_pztrad_2010, LangSE=clang_pztradsem_2010, 
         Math=cmath_pztrad_2010, MathSE=cmath_pztradsem_2010, 
         Read=cread_pztrad_2010, ReadSE=cread_pztradsem_2010, 
         Sci =cscic_pztrad_2010, SciSE =cscic_pztradsem_2010, 
         Lang1=clang_pztrad_2009, Lang1SE=clang_pztradsem_2009, 
         Math1=cmath_pztrad_2009, Math1SE=cmath_pztradsem_2009, 
         Read1=cread_pztrad_2009, Read1SE=cread_pztradsem_2009, 
         Sci1 =cscic_pztrad_2009, Sci1SE =cscic_pztradsem_2009, 
         Lang2=clang_pztrad_2008, Lang2SE=clang_pztradsem_2008, 
         Math2=cmath_pztrad_2008, Math2SE=cmath_pztradsem_2008, 
         Read2=cread_pztrad_2008, Read2SE=cread_pztradsem_2008, 
         Sci2 =cscic_pztrad_2008, Sci2SE =cscic_pztradsem_2008, 
         Lang3=clang_pztrad_2007, Lang3SE=clang_pztradsem_2007, 
         Math3=cmath_pztrad_2007, Math3SE=cmath_pztradsem_2007, 
         Read3=cread_pztrad_2007, Read3SE=cread_pztradsem_2007, 
         Sci3 =cscic_pztrad_2007, Sci3SE =cscic_pztradsem_2007, 
         HasTest=fs_cur, HasTest1=fs_lag1, HasTest01=fs_vamable, 
         PctEng=eng_tpct, PctSPED=sped_tpct, PctFRL=frl_tpct, 
         PctGift=gifted_tpct, PctA=asian_tpct, PctB=black_tpct,
         PctH=hisp_tpct, PctW=white_tpct, PctM=male_tpct, 
         AveLang1=clang_pztrad_2009_tavg, AveMath1=cmath_pztrad_2009_tavg,
         AveRead1=cread_pztrad_2009_tavg, AveSci1=cscic_pztrad_2009_tavg,
         AveLang0=clang_pztrad_2010_tavg, AveMath0=cmath_pztrad_2010_tavg,
         AveRead0=cread_pztrad_2010_tavg, AveSci0=cscic_pztrad_2010_tavg)

Y2 <- read_csv(file.path(loc,"StudentVAMData_Y2.csv")) %>%
  select(TID=utqid,CID=classid, dist, idc, syear, grdyr, cohort,SID=schoolid, sfrac, subject, 
         frac_t, iys_tccount, t_iyscount1, 
         eng, sped, frl, gifted, asian, black, hisp, white, male, 
         Lang=clang_pztrad_2011, LangSE=clang_pztradsem_2011, 
         Math=cmath_pztrad_2011, MathSE=cmath_pztradsem_2011, 
         Read=cread_pztrad_2011, ReadSE=cread_pztradsem_2011, 
         Sci =cscic_pztrad_2011, SciSE =cscic_pztradsem_2011, 
         Lang1=clang_pztrad_2010, Lang1SE=clang_pztradsem_2010, 
         Math1=cmath_pztrad_2010, Math1SE=cmath_pztradsem_2010, 
         Read1=cread_pztrad_2010, Read1SE=cread_pztradsem_2010, 
         Sci1 =cscic_pztrad_2010, Sci1SE =cscic_pztradsem_2010, 
         Lang2=clang_pztrad_2009, Lang2SE=clang_pztradsem_2009, 
         Math2=cmath_pztrad_2009, Math2SE=cmath_pztradsem_2009, 
         Read2=cread_pztrad_2009, Read2SE=cread_pztradsem_2009, 
         Sci2 =cscic_pztrad_2009, Sci2SE =cscic_pztradsem_2009, 
         Lang3=clang_pztrad_2008, Lang3SE=clang_pztradsem_2008, 
         Math3=cmath_pztrad_2008, Math3SE=cmath_pztradsem_2008, 
         Read3=cread_pztrad_2008, Read3SE=cread_pztradsem_2008, 
         Sci3 =cscic_pztrad_2008, Sci3SE =cscic_pztradsem_2008, 
         HasTest=fs_cur, HasTest1=fs_lag1, HasTest01=fs_vamable, 
         PctEng=eng_tpct, PctSPED=sped_tpct, PctFRL=frl_tpct, 
         PctGift=gifted_tpct, PctA=asian_tpct, PctB=black_tpct, 
         PctH=hisp_tpct, PctW=white_tpct, PctM=male_tpct, 
         AveLang1=clang_pztrad_2010_tavg, AveMath1=cmath_pztrad_2010_tavg,
         AveRead1=cread_pztrad_2010_tavg, AveSci1 =cscic_pztrad_2010_tavg,
         AveLang0=clang_pztrad_2011_tavg, AveMath0=cmath_pztrad_2011_tavg,
         AveRead0=cread_pztrad_2011_tavg, AveSci0 =cscic_pztrad_2011_tavg)

dim(Y1)
dim(Y2)
dim(dat <- bind_rows(Y1,Y2) %>% 
      select(-CID) %>% 
      unique 
      )
# rm(Y1,Y2)
dat <- dat %>% gather(Subj, Out, Lang, Math, Read, Sci) %>%
  select(-ends_with("SE")) %>%
  filter(Subj %in% c("Lang","Math")) %>%
  mutate(Subj=recode(Subj,Lang="Language")) %>%
  filter(Subj==subject) %>% select(-subject)
summary(dat)
# dat %>% count(idc,Subj,syear) %>% filter(n>1)
# dat %>% filter(idc=="S00001438") %>% as.data.frame
```

```{r, cache=T, message=F, warning=F, include=F}
datimp <- mice::mice(dat , m=1, maxit=5, meth="pmm",seed=1)
dat <- complete(datimp)
```

```{r models, cache=T, include=F, message=F, warning=F}
mod <- lmer(Out~1+(0+frac_t|TID)+(1|idc)+(0+sfrac|SID:Subj)+(0+sfrac|SID)+(0+sfrac|SID:grdyr),dat)
# summary(mod)
# VarPer(mod)

mod2 <- update(mod, .~.+Lang1+Math1+Lang2+Math2+Read1+Sci1+Read2+Sci2)
mod3 <- update(mod, .~.+eng+sped+frl+gifted+asian+black+hisp+male+Lang1+Math1+Lang2+Math2+Read1+Sci1+Read2+Sci2)
```

```{r modFRL, cache=T, include=F}
modFRL <- lmer(Out~1+(0+frac_t|TID)+(1|idc)+(0+sfrac|SID:Subj)+(0+sfrac|SID)+(0+sfrac|SID:grdyr),
               dat, subset=frl==1)
modNFRL <- lmer(Out~1+(0+frac_t|TID)+(1|idc)+(0+sfrac|SID:Subj)+(0+sfrac|SID)+(0+sfrac|SID:grdyr),
               dat, subset=frl==0)
MathT <- dat %>% count(TID,I(Math1<0)) %>% filter(n>5) %>% .$TID %>% unique
# modLang <- lmer(Out~1+(0+I((Lang1<0)*frac_t)+I((Lang1>=0)*frac_t)|TID)+
#                  (1|idc)+(0+sfrac|SID:Subj)+(0+sfrac|SID)+(0+sfrac|SID:grdyr),
#                 dat, subset=TID %in% MathT)
# MathT <- dat %>% count(TID,I(Math1<0)) %>% filter(n>5) %>% .$TID %>% unique
# modMath <- lmer(Out~1+(0+I((Math1<0)*frac_t)+I((Math1>=0)*frac_t)|TID)+
#                  (1|idc)+(0+sfrac|SID:Subj)+(0+sfrac|SID)+(0+sfrac|SID:grdyr),
#                 dat, subset=TID %in% MathT)
```

## Introduction to Value-Added Models

- Very Brief History
- Why Value-Added (VA)
- VA Conceptual Assumptions
- Evidence on VA Assumptions

## Not Covered

- Statisical Models
- Details on statistical models/intricacies


## Very Brief History {.smaller }

- Process-Product/school effects research used gain scores to look at teacher effects
    - Generally found large classroom/teacher effects
- Growth in minimum competency testing and later proficiency testing made scores available
    - TN analyses statewide found teacher as most important contributor (Sanders and Rivers)
- Led to policy interest in measuring teacher quality using test scores

## UTQ Data - Pct Variances {.smaller}

```{r}
data.frame(Unconditional=VarPer(mod), 
           PriorTestCtrl=VarPer(mod2),
           FullCtrl=VarPer(mod3) ) %>% 
  rownames_to_column %>% 
  mutate(rowname=recode(rowname, idc="Student", TID="Teacher", `SID:grdyr`="SchoolByGrade", `SID:Subj`="SchoolBySubj", `SID`="School")) %>%
  kable %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```

## Why Value-Added (VA)

- Desire for "objective" measure of teacher quality
    - Test scores as natural metric in proficiency based system
    - Test scores make it cheap to estimate
- Desire for outcome measure of schools
- Policy relevance of test scores in current regimes
    - Teachers are important source of differences in scores

## Defining VA as a Construct

What are the assumptions being made by VA scores? What construct is being measured?

- Students are taught by teachers in schools 
- Students learn many subjects, sometimes from many teachers
- VA score are estimated as a single number conditional on known student characteristics from previous years

## VA as Causal Model

- Potential Outcomes Framework
    - Each student has potential outcome if they are taught by each teacher
    - Observed outcomes are non-random subset
- Assumptions (all conditional on covariates)
    - Outcomes are missing at random (missingness independent of value)
    - Students have single potential outcome in each classroom (SUTVA)
    - See Reardon and Raudenbush, 2009
    
## UTQ Data {.smaller #Tabup}

<style> 
  #Tabup  { 
    margin-top: -40px; 
  } 
</style>

```{r}
coef(summary(mod3)) %>% kable %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = F)
```

## VA Assumptions - Stability {.smaller}

- Newton, et al., 2010 - correlations 0.43-0.63 across years in Math and 0.34-0.48 in ELA
- Papay, 2011 - 0.57-0.71
- Goldhaber & Hansen, 2010 - 0.32, 0.55 in ELA/Math

## VA Assumptions - Affected by SES {.smaller}

- Newton, et al., 2010 - VA scores are generally correlated with student characteristics and incoming student achievement. Teachers receive higher scores in high track versus low track courses
- Loeb, et al., 2014 - correlations of across ELL/non-ELL is 0.6
- Fox 2015 - highly stable VA scores across subgroups (cor>0.9)

## UTQ Data - Subgroups 

```{r UTQsub, cache=T}
fn <- Out~1+(0+frac_t|TID)+(1|idc)+(0+sfrac|SID:Subj)+(0+sfrac|SID)+(0+sfrac|SID:grdyr)
tbl<-data.frame(LowMath=VarPer(lmer(fn,dat, subset=Math1<=mean(Math1))), 
                HiMath =VarPer(lmer(fn,dat, subset=Math1>=mean(Math1))),
                LowLang=VarPer(lmer(fn,dat, subset=Lang1<=mean(Lang1))),
                HiLang =VarPer(lmer(fn,dat, subset=Lang1>=mean(Lang1))),
                FRL    =VarPer(modFRL),
                NonFRL =VarPer(modNFRL) 
                )
tbl %>% 
  rownames_to_column %>% 
  mutate(rowname=recode(rowname, idc="Student", TID="Teacher", `SID:grdyr`="SchoolByGrade", `SID:Subj`="SchoolBySubj", `SID`="School")) %>%
  kable %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```



## VA Assumptions - Dependence on Test {.smaller}

- Lockwood, et al., 2007 - correlation of VA scores across equivalent tests is 0.32-0.4
- Papay 2011 - correlations across tests 0.15-0.58
- Fox 2015 - correlations across subject 0.7

## VA Assumptions - Generalizing Across Schools {.smaller}

- Angrist, et al 2017 - Lottery randomization used to show VA scores are biased across schools
- Raudenbush 2013 - more theoretical but argues VA scores cannot generalize across different contexts

## Take-aways {.smaller}

- VA scores seem to be useful, but they are full of error
    - Randomization experiments suggest little within-school error, but potentially large between school error, especially when schools differ widely on student characteristics
- Averaging scores across 3+ years makes them more stable
    - Scores can then be used to identify teacher quality, but mostly in the tails
- VA scores are only available for some teachers and for fewer if aggregating over years
- Provide little opportunity for guidance for teachers
